Screening Sinkhorn Algorithm for Regularized Optimal Transport

### 2019-05-21 tex/v3/section_5_numeric_expes.tex
update

We compare \emph{Screenkhorn} to Sinkhorn as implemented in Python Optimal Transport toolbox\footnote{\url{https://pot.readthedocs.io/en/stable/index.html}} on  a synthetic example. The dataset we use consists of source samples generated from a bi-dimensional gaussian mixture and target samples following the same distribution but with different gaussian means. We consider an unsupervised domain adaptation using optimal transport with entropic regularization.  Several settings are explored: different values of $\eta$, the regularization parameter, the allowed budget $\frac{n_b}{n} = \frac{m_b}{m} = $ ranging from $0.01$ to $0.99$, different values of $n$ and $m$ and, whether the cost matrix $C$ is normalized or not. We empirically measure  marginal violations as the norms $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$, running time expressed as $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$ and the relative divergence difference $\| \inr{C, P^\star} - \inr{C, P^{sc}}\|$ between \emph{Screenkhorn} and Sinkhorn.

@@

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=4.55cm]{./figs/norm_M_Mu_marginals_toy_n1000} ~\hfill~
		\includegraphics[width=4.55cm]{./figs/norm_M_Mu_marginals_toy_n1000} ~\hfill \\
		\includegraphics[width=4.55cm]{./figs/norm_M_time_toy_n1000} ~\hfill~
		\includegraphics[width=4.55cm]{./figs/norm_M_div_toy_n1000}
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. Top panel: marginal violations in relation with the budget of points. Bottom panel: (left) ratio of computation times    $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$ and, (right) relative divergence variation. The results are averaged over 10 trials.} 
		\label{fig:margin_expe}
\end{figure*}
%
\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ''high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge towards zero when increasing the budget of points. Remarkably marginal violations are almost negligible whatever the budget for high $\eta$.  According to computation gain, \emph{Screenkhorn} is, up to 2 times, faster than Sinkhorn at low budget $n_b/n$ while the reverse holds when $n_b/n$ gets close to 1.  Computational benefit of \emph{Screenkhorn} also depends on $\eta$ with appropriate values $\eta \leq 1$. Finally except for $\eta=0.1$ \emph{Screenkhorn} achieves a close divergence $\inr{C, P}$ compared to Sinkhorn.

>>>>

We compare \emph{Screenkhorn} to Sinkhorn as implemented in Python Optimal Transport toolbox\footnote{\url{https://pot.readthedocs.io/en/stable/index.html}} on  a synthetic example. The dataset we use consists of source samples generated from a bi-dimensional gaussian mixture and target samples following the same distribution but with different gaussian means. We consider an unsupervised domain adaptation using optimal transport with entropic regularization.  Several settings are explored: different values of $\eta$, the regularization parameter, the allowed budget $\frac{n_b}{n} = \frac{m_b}{m} = $ ranging from $0.01$ to $0.99$, different values of $n$ and $m$.
% and, whether the cost matrix $C$ is normalized or not.
 We empirically measure  marginal violations as the norms $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$, running time expressed as $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$ and the relative divergence difference $\| \inr{C, P^\star} - \inr{C, P^{sc}}\|$ between \emph{Screenkhorn} and Sinkhorn.

@@

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=4.55cm]{./figs/norm_M_Mu_marginals_toy_n1000} ~\hfill~
		\includegraphics[width=4.55cm]{./figs/norm_M_Mu_marginals_toy_n1000} ~\hfill \\
		\includegraphics[width=4.55cm]{./figs/norm_M_time_toy_n1000} ~\hfill~
		\includegraphics[width=4.55cm]{./figs/norm_M_div_toy_n1000}
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. Top panel: marginal violations in relation with the budget of points. Bottom panel: (left) ratio of computation times    $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$ and, (right) relative divergence variation. The results are averaged over 10 trials.} 
		\label{fig:margin_expe}
\end{figure*}
%
\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ''high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge towards zero when increasing the budget of points. Remarkably marginal violations are almost negligible whatever the budget for high $\eta$.  According to computation gain, \emph{Screenkhorn} is, up to 2 times, faster than Sinkhorn at low budget $n_b/n$ while the reverse holds when $n_b/n$ gets close to 1.  Computational benefit of \emph{Screenkhorn} also depends on $\eta$ with appropriate values $\eta \leq 1$. Finally except for $\eta=0.1$ \emph{Screenkhorn} achieves a close divergence $\inr{C, P}$ compared to Sinkhorn showing that our static screening test does not harm Sinkhorn divergence.
